---
title: "Project 1"
author: "Chris Cavan, Marshall Ndhlovu, Noah Maddio, Nathan Gottwals"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Processing

```{r Libraries, warning=FALSE,message=FALSE,error=FALSE, results='hide'}
rm(list = ls())
# libraries 
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggcorrplot)
library(gridExtra)
library(tidyr)
```
All r code and rmarkdown will be in the root directory with and all sensory data are to be in the directory Data/ and extracted into their respective sub directories as stated:\

**Work order**

+ Data/work order scrubbed.csv
  
**Active Power**

+ Data/Active Power
  
**Gearbox HS Bearing Temperature**
  
+ Data/Gearbox Bearing Temperature
  
**Gearbox IMS Bearing**
  
+ Data/Gearbox IMS Bearing
  
**Gearbox Oil Temperature**

+ Gearbox Oil Temperature
  
**Generator RPM**

+ Data/Gearbox RPM
  
**Hydraulic Pressure**

+ Data/Hydraulic Pressure
  
**Windspeed**

+ Data/Windspeed

#Reading in the data and cleaning
The data for the supplied sensor metrics are read in and the dates and time are converted to a usable date format. The data of the metrics are
then aggregated by the average in 15 minute intervals. The data frames are then joined by the Turbine and with the date and time of the averaged metric.


```{r data_processing, eval=FALSE}
#### ***Rmarkdown: Will not run when kniting unless you remove eval=FALSE ^


#####workorder------------
wo = read.csv("Data/work order scrubbed.csv")
# work order exploration and data cleaning
# cause_code all null
#table(wo$cause_code)
# start date is character
#convert start and finish columns to date objects
wo$wo_start_date = ymd_hms(wo$wo_start_date)
# end date is also character
#convert start and finish columns to date objects
wo$wo_finish_date = ymd_hms(wo$wo_finish_date)
#table(wo$cause_code)

# This first converts "null" into a NA character. Then by grouping of OrderNo, fills in NA's with existing values of component_type, within a OrderNo if they exist. Then distinct() is ran to keep only unique rows. This allows us to keep unique work order numbers with component types variable where availible. 
workorder = wo %>% 
    mutate(component_type = case_when(component_type == "null" ~ NA_character_,
                               TRUE ~ component_type)) %>% 
    group_by(OrderNo) %>% 
    fill(component_type, .direction = 'up') %>% 
    fill(component_type, .direction = 'down') %>% distinct(OrderNo, .keep_all = TRUE)
# compared to this which just keeps distinct OrderNo regaurdless of other variables, causing use to lose most of the component_type 
# test = distinct(wo, OrderNo, .keep_all = TRUE)

str(workorder)

#### fault code---------------- 
fault_files = list.files("Data/Fault Codes", pattern="*.csv")
fault_list = list()
for (i in 1:length(fault_files)) {
  if (file.size(paste0("Data/Fault Codes/", fault_files[i])) >1){
    fault_list = c(fault_list, fault_files[i])
  }
}
fault = read.csv(paste0("Data/Fault Codes/", fault_list[1]), header = FALSE)
for (i in 2:length(fault_list)){
  new = read.csv(paste0("Data/Fault Codes/", fault_list[i]), header = FALSE)
  fault = rbind(fault, new)
}
#convert Datetime from char to date
fault$V2 <- ymd_hms(fault$V2)
#renaming column headers
names(fault) <- c("location_id", "DateTime", "Date", "FaultCode", "StatusCode", "Description", "Type")
fault <- fault %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
fault$round_date <- lubridate::round_date(fault$DateTime, "15 minutes")




#### Gearbox HS Bearing Temperature---------------- 
bearing_files = list.files("Data/Gearbox Bearing Temperature", pattern="*.csv")
bearing_list = list()
for (i in 1:length(bearing_files)) {
  if (file.size(paste0("Data/Gearbox Bearing Temperature/", bearing_files[i])) >1){
    bearing_list = c(bearing_list, bearing_files[i])
  }
}
bearing = read.csv(paste0("Data/Gearbox Bearing Temperature/", bearing_list[1]), header = FALSE)
for (i in 2:length(bearing_list)){
  new = read.csv(paste0("Data/Gearbox Bearing Temperature/", bearing_list[i]), header = FALSE)
  bearing = rbind(bearing, new)
}
#convert Datetime from char to date
bearing$V2 <- ymd_hms(bearing$V2)
#renaming column headers
names(bearing) <- c("location_id", "DateTime", "Date", "Temperature", "Bearing DE")
# Sorting data by turbine, then the DateTime. Then deleting rows with duplicate start times by turbine
bearing <- bearing %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
bearing$round_date <- lubridate::round_date(bearing$DateTime, "15 minutes")
#adding mean of the 15 minute intervals and dropping columns
bearing <- bearing %>% group_by(location_id, round_date) %>% summarise(mean_temp = mean(Temperature), .groups = 'drop') %>% as.data.frame()



#### Gearbox IMS Bearing Temperature---------------- 
ims_files = list.files("Data/Gearbox IMS Bearing", pattern="*.csv")
ims_list = list()
for (i in 1:length(ims_files)) {
  if (file.size(paste0("Data/Gearbox IMS Bearing/", ims_files[i])) >1){
    ims_list = c(ims_list, ims_files[i])
  }
}
ims = read.csv(paste0("Data/Gearbox IMS Bearing/", ims_list[1]), header = FALSE)
for (i in 2:length(ims_list)){
  new = read.csv(paste0("Data/Gearbox IMS Bearing/", ims_list[i]), header = FALSE)
  ims = rbind(ims, new)
}
#convert Datetime from char to date
ims$V2 <- ymd_hms(ims$V2)
#renaming column headers
names(ims) <- c("location_id", "DateTime", "Date", "Temperature", "ims")
# Sorting data by turbine, then the DateTime. Then deleting rows with duplicate start times by turbine
ims <- ims %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
ims$round_date <- lubridate::round_date(ims$DateTime, "15 minutes")
#adding mean ims bearing temp of the 15 minute intervals and dropping columns
ims <- ims %>% group_by(location_id, round_date) %>% summarise(mean_ims = mean(Temperature), .groups = 'drop') %>% as.data.frame()



#### Gearbox Oil Temperature---------------- 
oil_files = list.files("Data/Gearbox Oil Temperature", pattern="*.csv")
oil_list = list()
for (i in 1:length(oil_files)) {
  if (file.size(paste0("Data/Gearbox Oil Temperature/", oil_files[i])) >1){
    oil_list = c(oil_list, oil_files[i])
  }
}
oil = read.csv(paste0("Data/Gearbox Oil Temperature/", oil_list[1]), header = FALSE)
for (i in 2:length(oil_list)){
  new = read.csv(paste0("Data/Gearbox Oil Temperature/", oil_list[i]), header = FALSE)
  oil = rbind(oil, new)
}
#convert Datetime from char to date
oil$V2 <- ymd_hms(oil$V2)
#renaming column headers
names(oil) <- c("location_id", "DateTime", "Date", "Temperature", "oil")
# Sorting data by turbine, then the DateTime. Then deleting rows with duplicate start times by turbine
oil <- oil %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
oil$round_date <- lubridate::round_date(oil$DateTime, "15 minutes")
#adding mean oil temp of the 15 minute intervals and dropping columns
oil <- oil %>% group_by(location_id, round_date) %>% summarise(mean_oil = mean(Temperature), .groups = 'drop') %>% as.data.frame()



#### Generator RPM---------------- 
rpm_files = list.files("Data/Generator RPM", pattern="*.csv")
rpm_list = list()
for (i in 1:length(rpm_files)) {
  if (file.size(paste0("Data/Generator RPM/", rpm_files[i])) >1){
    rpm_list = c(rpm_list, rpm_files[i])
  }
}
rpm = read.csv(paste0("Data/Generator RPM/", rpm_list[1]), header = FALSE)
for (i in 2:length(rpm_list)){
  new = read.csv(paste0("Data/Generator RPM/", rpm_list[i]), header = FALSE)
  rpm = rbind(rpm, new)
}
#convert Datetime from char to date
rpm$V2 <- ymd_hms(rpm$V2)
#renaming column headers
names(rpm) <- c("location_id", "DateTime", "Date", "RPM", "Generator rpm")
# Sorting data by turbine, then the DateTime. Then deleting rows with duplicate start times by turbine
rpm <- rpm %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
rpm$round_date <- lubridate::round_date(rpm$DateTime, "15 minutes")
#adding mean rpm of the 15 minute intervals and dropping columns
rpm <- rpm %>% group_by(location_id, round_date) %>% summarise(mean_rpm = mean(RPM), .groups = 'drop') %>% as.data.frame()



#### Hydraulic Pressure---------------- 
hydraulic_files = list.files("Data/Hydraulic Pressure", pattern="*.csv")
hydraulic_list = list()
for (i in 1:length(rpm_files)) {
  if (file.size(paste0("Data/Hydraulic Pressure/", hydraulic_files[i])) >1){
    hydraulic_list = c(hydraulic_list, hydraulic_files[i])
  }
}
hydraulic = read.csv(paste0("Data/Hydraulic Pressure/", hydraulic_list[1]), header = FALSE)
for (i in 2:length(hydraulic_list)){
  new = read.csv(paste0("Data/Hydraulic Pressure/", hydraulic_list[i]), header = FALSE)
  hydraulic = rbind(hydraulic, new)
}
#convert Datetime from char to date
hydraulic$V2 <- ymd_hms(hydraulic$V2)
#renaming column headers
names(hydraulic) <- c("location_id", "DateTime", "Date", "Pressure", "Hydraulic Pressure")
# Sorting data by turbine, then the DateTime. Then deleting rows with duplicate start times by turbine
hydraulic <- hydraulic %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
hydraulic$round_date <- lubridate::round_date(hydraulic$DateTime, "15 minutes")
#adding mean rpm of the 15 minute intervals and dropping columns
hydraulic <- hydraulic %>% group_by(location_id, round_date) %>% summarise(mean_pressure = mean(Pressure), .groups = 'drop') %>% as.data.frame()



#### active power-------------- 
active_files = list.files("Data/Active Power", pattern="*.csv")
active_list = list()
for (i in 1:length(active_files)) {
  if (file.size(paste0("Data/Active Power/", active_files[i])) >1){
    active_list = c(active_list, active_files[i])
  }
}
active = read.csv(paste0("Data/Active Power/", active_list[1]), header = FALSE)
for (i in 2:length(active_list)){
  new = read.csv(paste0("Data/Active Power/", active_list[i]), header = FALSE)
  active = rbind(active, new)
}
#convert Datetime from char to date
active$V2 <- ymd_hms(active$V2)
#renaming column headers
names(active) <- c("location_id", "DateTime", "Date", "power", "Active power")
# sorting data by turbine, then the DateTime.
active <- active %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
active$round_date <- lubridate::round_date(active$DateTime, "15 minutes")
#adding mean power of the 15 minute intervals and dropping columns
active <-active %>% group_by(location_id, round_date) %>% summarise(mean_power = mean(power), .groups = 'drop') %>% as.data.frame()



#### Ambient Temperature-------------- Needs work*****
ambient_files = list.files("Data/ambient_temperature", pattern="*.csv")
ambient_list = list()
for (i in 1:length(ambient_files)) {
  if (file.size(paste0("Data/ambient_temperature/", ambient_files[i])) >1){
    ambient_list = c(ambient_list, ambient_files[i])
  }
}
ambient = read.csv(paste0("Data/ambient_temperature/", ambient_list[1]), header = FALSE)
for (i in 2:length(ambient_list)){
  new = read.csv(paste0("Data/ambient_temperature/", ambient_list[i]), header = FALSE)
  ambient = rbind(ambient, new)
}
#convert Datetime from char to date
ambient$V2 <- ymd_hms(ambient$V2)
#renaming column headers
names(ambient) <- c("location_id", "DateTime", "Date", "ambient_temp", "Ambient Temperature")
# sorting data by turbine, then the DateTime.
ambient <- ambient %>% 
  arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
ambient$round_date <- lubridate::round_date(ambient$DateTime, "15 minutes")
#adding mean power of the 15 minute intervals and dropping columns
ambient <-ambient %>% group_by(location_id, round_date) %>% summarise(mean_ambient = mean(ambient_temp), .groups = 'drop') %>% as.data.frame

#### Wind Speed-------------- 
wind_files = list.files("Data/Windspeed", pattern="*.csv")
wind_list = list()
for (i in 1:length(wind_files)) {
  if (file.size(paste0("Data/Windspeed/", wind_files[i])) >1){
    wind_list = c(wind_list, wind_files[i])
  }
}
wind = read.csv(paste0("Data/Windspeed/", wind_list[1]), header = FALSE)
for (i in 2:length(wind_list)){
  new = read.csv(paste0("Data/Windspeed/", wind_list[i]), header = FALSE)
  wind = rbind(wind, new)
}
#convert Datetime from char to date
wind$V2 <- ymd_hms(wind$V2)
#renaming column headers
names(wind) <- c("location_id", "DateTime", "Date", "windspeed", "Wind Speed")
# sorting data by turbine, then the DateTime.
wind <- wind %>% arrange(location_id, DateTime) %>% distinct()
#Adding a variable round_date to round times to nearest 15 minute intervals
wind$round_date <- lubridate::round_date(wind$DateTime, "15 minutes")
#adding mean power of the 15 minute intervals and dropping columns
wind <-wind %>% group_by(location_id, round_date) %>% summarise(mean_windspeed = mean(windspeed), .groups = 'drop') %>% as.data.frame()



####Joining datasets------
fully_joined <- full_join(fault, bearing, by=c('location_id', 'round_date')) %>%
  full_join(., wind, c('location_id', 'round_date')) %>%
  full_join(., hydraulic, c('location_id', 'round_date')) %>%
  full_join(., rpm, c('location_id', 'round_date')) %>%
  full_join(., oil, c('location_id', 'round_date')) %>%
  full_join(., ims, c('location_id', 'round_date')) %>%
  full_join(., active, c('location_id', 'round_date'))%>%
  full_join(., ambient, by=c('location_id', 'round_date'))

# Removing unused columns/duplicates
fully_joined <- subset(fully_joined, select = -c(DateTime, Date, StatusCode, Type, Date))

# Renaming columns
names(fully_joined) <- c("Turbine", "FaultCode","Discription", "Timestamp", "Bearing_Temp", "Windspeed", "Hydraulic_Pressure", "Generator_RPM", "Oil_Temp", "Gearbox_IMS", "Active_Power", "Ambient_Temp")


# Output Dataset not imputed
write.csv(fully_joined, "final_data-v2.csv", row.names=FALSE)



#Impute Dataset
imputed_dataset = fully_joined
#Replace all NA values in numeric columns with average column value.
imputed_dataset$Bearing_Temp[is.na(imputed_dataset$Bearing_Temp)] <- mean(imputed_dataset$Bearing_Temp,na.rm=TRUE)
imputed_dataset$Windspeed[is.na(imputed_dataset$Windspeed)] <- mean(imputed_dataset$Windspeed,na.rm=TRUE)
imputed_dataset$Hydraulic_Pressure[is.na(imputed_dataset$Hydraulic_Pressure)] <- mean(imputed_dataset$Hydraulic_Pressure,na.rm=TRUE)
imputed_dataset$Generator_RPM[is.na(imputed_dataset$Generator_RPM)] <- mean(imputed_dataset$Generator_RPM,na.rm=TRUE)
imputed_dataset$Oil_Temp[is.na(imputed_dataset$Oil_Temp)] <- mean(imputed_dataset$Oil_Temp,na.rm=TRUE)
imputed_dataset$Gearbox_IMS[is.na(imputed_dataset$Gearbox_IMS)] <- mean(imputed_dataset$Gearbox_IMS,na.rm=TRUE)
imputed_dataset$Active_Power[is.na(imputed_dataset$Active_Power)] <- mean(imputed_dataset$Active_Power,na.rm=TRUE)
imputed_dataset$Ambient_Temp[is.na(imputed_dataset$Ambient_Temp)] <- mean(imputed_dataset$Ambient_Temp,na.rm=TRUE)

# binary variable FaultOccurred. If a fault code is present in the time interval, Yes, if not,No.
imputed_dataset$FaultOccurred <-with(imputed_dataset, ifelse(is.na(db$FaultCode),"Yes","No"))
as.factor(imputed_dataset$FaultOccurred)

# Output Dataset impouted
write.csv(imputed_dataset, "imputed_final_data.csv", row.names=FALSE)


```

## Matrix

```{r matrix, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='all'}
#Read in imputed dataset
db <- read.csv("Data/imputed_final_data.csv")

library(ggcorrplot)
numeric <- dplyr::select_if(db, is.numeric)
numeric <- subset(numeric, select = -c(FaultCode)) # removing Fault code as it is a unique set
r <- cor(numeric, use="pairwise.complete.obs")
round(r,2)
ggcorrplot(r) + labs(title = "Correlation matrix")
#Notes:
# We can see that the Gearbox oil temp and IMS temperature are directly related with correlation of .93
# We can see that Active Power is also directly related to the Generator RPM
# We can see that Wind Speed weakly inversely related to Hydraulic Pressure, Generator RPM, and Gearbox Oil temperature

```

## Multivariate
```{r multivariate, figures-side, fig.show="hold", out.width="100%", fig.align="center"}


p1 <- ggplot(db, aes(x=Fault_Occurred, y=Bearing_Temp)) + 
  geom_boxplot()
p2 <- ggplot(db, aes(x=Fault_Occurred, y=Windspeed)) + 
  geom_boxplot()
p3 <- ggplot(db, aes(x=Fault_Occurred, y=Hydraulic_Pressure)) + 
  geom_boxplot()
p4 <- ggplot(db, aes(x=Fault_Occurred, y=Generator_RPM)) + 
  geom_boxplot()
p5 <- ggplot(db, aes(x=Fault_Occurred, y=Oil_Temp)) + 
  geom_boxplot()
p6 <- ggplot(db, aes(x=Fault_Occurred, y=Gearbox_IMS)) + 
  geom_boxplot()
p7 <- ggplot(db, aes(x=Fault_Occurred, y=Active_Power)) + 
  geom_boxplot()
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=2)

ggplot(db,
       aes(x=Generator_RPM,
           y=Active_Power, 
           group=Fault_Occurred, 
           color=Fault_Occurred))+
  ggtitle("Mean Active Power Vs Mean RPM")+
  xlab("Mean Genorator RPM")+
  ylab("Mean Active Power")+
  geom_point()

ggplot(db,
       aes(x=Oil_Temp,
           y=Gearbox_IMS, 
           group=Fault_Occurred, 
           color=Fault_Occurred))+
  ggtitle("Mean Gearbox Oil Temperature Vs Mean Gearbox IMS Bearing Temperature")+
  xlab("Mean Gearbox Oil Temperature")+
  ylab("Mean Gearbox IMS Bearing temperature")+
  coord_cartesian(xlim = c(20,75), ylim = c(20,75))+
  geom_point()

ggplot(db,
       aes(x=Windspeed,
           y=Hydraulic_Pressure, 
           group=Fault_Occurred, 
           color=Fault_Occurred))+
  ggtitle("Mean Wind Speed Vs Mean Hydraulic Pressure")+
  xlab("Mean Wind Speed")+
  ylab("Mean Hydraulic Pressure")+
  geom_point()

ggplot(db,
       aes(x=Windspeed,
           y=Active_Power, 
           group=Fault_Occurred, 
           color=Fault_Occurred))+
  ggtitle("Mean Wind Speed Vs Mean Active Power")+
  xlab("Mean Wind Speed")+
  ylab("Mean Active Power")+
  geom_point()

ggplot(db, aes(Active_Power)) +
geom_histogram(aes(fill = Fault_Occurred), color = "black", binwidth = 50) +
ggtitle("Histogram of Active Power by Fault Occurred") +
xlab("Power Output") +
ylab("Number of Obvs.")

oilPow = subset(db, select = c(Oil_Temp, Active_Power, Fault_Occurred))
oilPow = na.omit(oilPow)

plot(oilPow$Active_Power, oilPow$Oil_Temp, main="Active Power vs. Oil Temp",
     xlab="Active Power", ylab="Oil Temp", pch=19, col = factor(oilPow$Fault_Occurred))



```
